---
title: "MPG Per Transmission"
author: "EGB"
date: "June 26, 2016"
output: html_document
---

The following analysis looks into the affect of miles per gallon (MPG) based on type of transmission.  We look at the **mtcars** dataset and compare the impact on MPG from a manual transmission versus an automatic one.  By doing the analysis below, we find that on average a *MANUAL* transmission will give you a higher MPG rate.

###Getting Data

```{r}
library(datasets)
data("mtcars")
```

###Exploratory Data Analysis

```{r}
str(mtcars)
head(mtcars)
```

###Standard Linear Model (LM)

```{r}
CR <- lm(mpg ~ am, data = mtcars)
summary(CR)
```

We see that transmission is very significant in determining MPG. If you have an automatic transmission, your average MPG is 17.147 while a manual transmission your average is +7.245 more making it 24.392 MPG.

```{r, echo=FALSE}
plot(mtcars$am,mtcars$mpg,pch=19,col="blue",xlab="Transmission: 0-Automatic, 1-Manual", ylab="MPG", main = "Fit with \
Standard LM")
abline(CR, col = "red", lwd = 3)
```

From the plot, we have further support that manual transmissions tend to have higher MPG's.  

###Poisson General Linear Model (GLM)

```{r, echo=FALSE, warning=FALSE}
plot(mtcars$am,mtcars$mpg,pch=19,col="blue",xlab="Transmission: 0-Automatic, 1-Manual", ylab="MPG", main = "Fit with \
Standard LM and Poisson GLM")
glm1 = glm(mtcars$mpg ~ mtcars$am,family="poisson")
abline(CR,col="red",lwd=3); lines(mtcars$am,glm1$fitted,col="blue",lwd=3)
legend("topleft", lty = 1, col = c("blue", "red"), legend = c("Poisson GLM", "Standard LM"))
```

Running a Poisson GLM did not result in a better fit.  It is the same as with a standard linear model.

###Residuals and Residual Plots
####Residuals

```{r}
n <- length(mtcars$mpg)
e <- resid(CR)
yhat <- predict(CR)
max(abs(e -(mtcars$mpg - yhat)))
max(abs(e - (mtcars$mpg  - coef(CR)[1] - coef(CR)[2] * mtcars$am)))
```

####Residual Plots

```{r, echo=FALSE, fig.align="center"}
par(mfrow = c(1,2), mar = c(4,4,6,1), oma = c(0,0,2,0))
plot(mtcars$am, mtcars$mpg,  
     xlab = "Transmission", 
     ylab = "MPG", 
     bg = "lightblue", 
     col = "black", cex = 2, pch = 21,frame = FALSE,
     main = "Residuals")
abline(CR, lwd = 2)
for (i in 1 : n) 
  lines(c(mtcars$am[i], mtcars$am[i]), c(mtcars$mpg [i], yhat[i]), col = "red" , lwd = 2)
 
plot(mtcars$am, e,  
     xlab = "Transmission", 
     ylab = "MPG Residuals", 
     bg = "lightblue", 
     col = "black", cex = 2, pch = 21,frame = FALSE,
     main = "Residuals versus X")
abline(h = 0, lwd = 2)
for (i in 1 : n) 
  lines(c(mtcars$am[i], mtcars$am[i]), c(e[i], 0), col = "red" , lwd = 2)
mtext("Residual Plots1", outer = TRUE)
```

####Better Residual Plots
#####Standard Linear Model

```{r, echo=FALSE, fig.align="center"}
par(mfrow= c(2,2))
plot(CR)
```

#####Poisson GLM

```{r, echo=FALSE, fig.align="center"}
par(mfrow= c(2,2))
plot(glm1)
```

###Confidence Intervals
```{r, warning=FALSE}
confint(CR)
confint(glm1)
```

From this, we see that although the fit may not show a difference, the confidence of the Poisson GLM is greater than the standard linear model.